---
title: "Untitled"
output: html_document
date: "2025-05-23"
---

### importing packages

```{r}
library(tidyverse)
library(ggrepel)
library(geodata)
library(sf)
library(geosphere)
library(cartogram)
library(readxl)
library(magick)
library(viridis)
library(ggtext)
```

### loading in data
```{r}

#listing female files files
file_paths <- list.files("new_data", pattern = "^f_\\d+-\\d+\\.csv$", full.names = TRUE)

#reading and cleaning each file
data_list <- lapply(file_paths, function(path) {
  df <- read_csv2(path, locale = locale(encoding = "latin1"))

  # Get x and y from filename
  x_y <- str_match(basename(path), "f_(\\d+)-(\\d+)\\.csv")[,2:3]
  colnames(df)[5:14] <- as.character(seq(as.integer(x_y[1]), as.integer(x_y[2])))

  # Rename and clean 'orig' and 'dest' columns
  df <- df %>%
    mutate(
    orig = sub("^Fra ", "", orig),
    dest = sub("^Til ", "", dest)
    )

  return(df)
})

#creating female df
combined_df_female <- data_list[[1]]

#binding only the numeric columns (last 10) from the rest
for (i in 2:length(data_list)) {
  numeric_cols <- data_list[[i]] %>% select((ncol(.) - 9):ncol(.))
  combined_df_female<- bind_cols(combined_df_female, numeric_cols)
}

### doing the same for the male data

# listing all male files
file_paths <- list.files("new_data", pattern = "^m_\\d+-\\d+\\.csv$", full.names = TRUE)

#reading and cleaning each file
data_list <- lapply(file_paths, function(path) {
  df <- read_csv2(path, locale = locale(encoding = "latin1"))

  #extracting age
  x_y <- str_match(basename(path), "m_(\\d+)-(\\d+)\\.csv")[,2:3]
  colnames(df)[5:14] <- as.character(seq(as.integer(x_y[1]), as.integer(x_y[2])))

  #renaming and cleaning 'orig' and 'dest' columns
  df <- df %>%
    mutate(
      orig = sub("^Fra ", "", orig),
      dest = sub("^Til ", "", dest)
    )

  return(df)
})

#saving male 
combined_df_male <- data_list[[1]]

# Bind only the numeric columns (last 10) from the rest
for (i in 2:length(data_list)) {
  numeric_cols <- data_list[[i]] %>% select((ncol(.) - 9):ncol(.))
  combined_df_male  <- bind_cols(combined_df_male, numeric_cols)
}

#creating gender columns
combined_df_male <- combined_df_male %>% mutate(sex = 1)
combined_df_female <- combined_df_female %>% mutate(sex = 0)
names(combined_df_female) <- names(combined_df_male)  # ensure columns match

combined_df <- rbind(combined_df_male, combined_df_female)

#combining the two
df <- combined_df %>% select(-'Mænd' & -'2024')


#### note: this process was necessary as statistikbanken has a limit to how much data can be downloaded at once - therefore, male and female were downladed separately in 10 age groups each - these are then combined in this code chunk to create one singular dataframe.
```


### cleaning data
```{r}
# Normalize city names: convert "Århus" to "Aarhus"
df$orig <- recode(df$orig, "Århus" = "Aarhus")
df$dest <- recode(df$dest, "Århus" = "Aarhus")


# adding total column
df <- df %>%
  mutate(total = rowSums(select(., matches("^\\d+$"))))

```


### loading in municipalities and transforming to sf
```{r}
munic <- gadm(country = "DNK", level = 2, path = ".")

# Convert to sf
munic_sf <- st_as_sf(munic)

# Get centroids and coordinates
centroids <- st_centroid(munic_sf)
```

# extract coordinates and attach muni name from NAME_2; calculating centroids and extract coordinates safely
```{r}
centroids_coords <- munic_sf %>%
  st_centroid() %>%
  mutate(
    lon = st_coordinates(.)[, 1],
    lat = st_coordinates(.)[, 2],
    muni = NAME_2  # explicitly create a 'muni' column
  ) %>%
  st_drop_geometry()  # drop geometry so it's a plain tibble

centroids_coords$muni <- recode(centroids_coords$muni, "Århus" = "Aarhus")

```

# calculating global maximum of movers (ved ik om jeg skal bruge)
```{r}
# # Calculate the global maximum number of movers
# global_max_movers <- df %>%
#   group_by(orig, dest) %>%
#   summarise(movers = sum(total), .groups = "drop") %>%
#   summarise(max_movers = max(movers)) %>%
#   pull(max_movers)



```

# plotting flow line maps
```{r}


# ─── Define age groups ─────────────────────────────────────────────────────────
age_groups <- list(
  "0-17"   = 0:17,
  "18-29"  = 18:29,
  "30-44"  = 30:44,
  "45-65"  = 45:65,
  "66-99"  = 66:99
)

# ─── Define top cities ─────────────────────────────────────────────────────────
top5_cities <- c("København", "Aarhus", "Odense", "Aalborg")

# ─── Calculate global max movers for consistent alpha & linewidth scaling ──────
all_age_cols <- paste0(0:99)
df$total_movers <- rowSums(df[all_age_cols], na.rm = TRUE)
global_max_movers_log <- log1p(max(df$total_movers, na.rm = TRUE))

# ─── Loop over each age group ──────────────────────────────────────────────────
for (age_label in names(age_groups)) {
  age_range <- age_groups[[age_label]]
  age_cols <- paste0(age_range)

  # 1. Aggregate by age group
  df_age <- df %>%
    mutate(total = rowSums(select(., all_of(age_cols)), na.rm = TRUE))

  # 2. Loop over each city
  for (city in top5_cities) {

    flow_city <- df_age %>%
      filter(dest == city, orig != dest) %>%
      group_by(orig, dest) %>%
      summarise(movers = sum(total), .groups = "drop") %>%
      left_join(centroids_coords, by = c("orig" = "muni")) %>%
      rename(lon_o = lon, lat_o = lat) %>%
      left_join(centroids_coords, by = c("dest" = "muni")) %>%
      rename(lon_d = lon, lat_d = lat) %>%
      filter(!is.na(lon_o), !is.na(lon_d)) %>%
      mutate(
        movers_log = log1p(movers),
        alpha_scaled = movers_log / global_max_movers_log
      )

    # 5. Select top origins for labeling
    top_origins <- flow_city %>%
      slice_max(order_by = movers, n = 10)

    # 6. Plot
    p <- ggplot() +
      geom_sf(data = munic_sf, fill = "lightgreen", color = "white", size = 0.2) +
      geom_segment(
        data = flow_city,
        aes(x = lon_o, y = lat_o, xend = lon_d, yend = lat_d,
            linewidth = movers_log, alpha = alpha_scaled),
        color = "steelblue"
      ) +
      geom_text_repel(
        data = top_origins,
        aes(x = lon_o, y = lat_o, label = orig),
        size = 3,
        color = "black",
        max.overlaps = Inf
      ) +
      scale_linewidth_continuous(
        range = c(0.3, 2.5),  # Adjust thickness as needed
        limits = c(0, global_max_movers_log),
        name = "Log-scaled movers"
      ) +
      scale_alpha_continuous(
        name = "Relative flow (log scale)",
        limits = c(0, 1)
      ) +
      guides(
        alpha = guide_legend(order = 1),
        linewidth = guide_legend(order = 2)
      ) +
      theme_minimal(base_size = 14) +
      theme(
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        panel.background = element_rect(fill = "lightblue", color = NA),
        plot.background = element_rect(fill = "lightblue", color = NA),
        plot.subtitle = element_markdown(size = 12),
        plot.title = element_markdown(size = 14)
      ) +
      labs(
        title = paste0("Migration Flows into <b>", city, "</b> 2024"),
        subtitle = paste0(
          "<span style='color:darkred;'><b>Age group: ", age_label,
          "</b></span> | Arrows show width & opacity ∝ movers"
        ),
        x = 'Longitude',
        y = 'Latitude'
      )

    # Save the plot
    filename <- paste0("out/migration_flows_", city, "_age_", gsub("-", "_", age_label), ".png")
    ggsave(filename = filename, plot = p, width = 8, height = 6, dpi = 300)

    print(p)
  }
}
## note: a radius circle was initially created, but we elected not to include it in our final product - the code still remains but is crossed out for curious minds who want to see it.
```

# creating gif
```{r}
pacman::p_load('magick')
library(magick)

#defining parameters
age_labels <- names(age_groups)
output_dir <- "out"
gif_dir <- "gifs"
dir.create(gif_dir, showWarnings = FALSE)

for (city in top5_cities) {
  # Generate file paths for this city's images
  image_files <- paste0(output_dir, "/migration_flows_", city, "_age_", gsub("-", "_", age_labels), ".png")
  
  # Read and combine images
  images <- image_read(image_files)
  animation <- image_animate(image_join(images), fps = 0.5 )  # Adjust fps as desired
  
  # Save gif
  gif_filename <- file.path(gif_dir, paste0("migration_flows_", city, ".gif"))
  image_write(animation, gif_filename)
}

```

# calculating and visualising mean kilometers traveled per relocation by age and age groups
```{r}
# Prepare a dataframe for storing results
age_distances_all <- data.frame()

#loop through each age (0 to 99)
for (age in 0:99) {
  
  # Calculate total movers for this specific age
  df_age <- df %>%
    filter(!is.na(.[[as.character(age)]])) %>%
    mutate(total = .[[as.character(age)]]) %>%
    filter(total > 0)  # Remove rows with zero movers
  
  # Calculate distances for each movement
  distances <- df_age %>%
    left_join(centroids_coords, by = c("orig" = "muni")) %>%
    rename(lon_o = lon, lat_o = lat) %>%
    left_join(centroids_coords, by = c("dest" = "muni")) %>%
    rename(lon_d = lon, lat_d = lat) %>%
    filter(!is.na(lon_o), !is.na(lon_d)) %>%
    mutate(distance_km = distHaversine(cbind(lon_o, lat_o), cbind(lon_d, lat_d)) / 1000) %>%
    summarise(
      mean_distance = mean(distance_km),
      sd_distance = sd(distance_km),
      n = n()
    ) %>%
    mutate(
      age = age,
      se = sd_distance / sqrt(n),  # Standard error
      ci_lower = mean_distance - 1.96 * se,
      ci_upper = mean_distance + 1.96 * se
    )
  
  # Append to the main dataframe
  age_distances_all <- bind_rows(age_distances_all, distances)
}

age_distances_all <- age_distances_all %>%
  mutate(age_group = case_when(
    age <= 17 ~ "0–17",
    age <= 29 ~ "18–29",
    age <= 44 ~ "30–44",
    age <= 65 ~ "45–65",
    TRUE ~ "66–99"
  ))


# Create segment dataframe for colored line segments
segment_data <- age_distances_all %>%
  arrange(age) %>%
  mutate(
    xend = lead(age),
    yend = lead(mean_distance),
    group_end = lead(age_group)
  ) %>%
  filter(!is.na(xend))


# Plot
ggplot() +
  # Confidence interval ribbon in grey
  geom_ribbon(data = age_distances_all, aes(x = age, ymin = ci_lower, ymax = ci_upper), 
              fill = "grey80", alpha = 0.3) +
  
  # Base grey line to maintain continuity
  geom_line(data = age_distances_all, aes(x = age, y = mean_distance), color = "grey80", size = 1.2) +
  
  # Colored segments by age group
  geom_segment(data = segment_data, aes(x = age, y = mean_distance, 
                                        xend = xend, yend = yend, color = age_group), size = 1.5) +
  
  # Points colored by age group
  geom_point(data = age_distances_all, aes(x = age, y = mean_distance, color = age_group), size = 2) +
  
  labs(
    title = "Mean Kilometers Traveled per Age (All Municipalities)",
    x = "Age",
    y = "Mean Distance (km)",
    color = "Age Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(breaks = seq(0, 99, 5)) +
  scale_y_continuous(breaks = seq(20, 100, 10)) +
  scale_color_manual(values = c(
    "0–17" = "#E41A1C",
    "18–29" = "#377EB8",
    "30–44" = "#4DAF4A",
    "45–65" = "#984EA3",
    "66–99" = "#FF7F00"
  ))

# Save the plot
ggsave("out/mean_distance_by_age_colored.png", width = 10, height = 6, dpi = 300)
```

## calculating and visualizing total relocations per age and age group
```{r}
# Prepare a dataframe for storing results
age_relocations <- data.frame()

# Loop through each age (0 to 99)
for (age in 0:99) {
  total_relocations <- df %>%
    filter(!is.na(.[[as.character(age)]])) %>%
    summarise(total_relocations = sum(.[[as.character(age)]], na.rm = TRUE)) %>%
    mutate(age = age)
  
  age_relocations <- bind_rows(age_relocations, total_relocations)
}

# Create age group column
age_relocations <- age_relocations %>%
  mutate(age_group = case_when(
    age <= 17 ~ "0–17",
    age <= 29 ~ "18–29",
    age <= 44 ~ "30–44",
    age <= 65 ~ "45–65",
    TRUE ~ "66–99"
  ))

# Create a segment dataframe for colored line segments
segment_data <- age_relocations %>%
  arrange(age) %>%
  mutate(
    xend = lead(age),
    yend = lead(total_relocations),
    group_end = lead(age_group)
  ) %>%
  filter(!is.na(xend))  # Remove last row which has NA in lead

# Plot continuous line in grey and overlay colored segments
ggplot() +
  geom_line(data = age_relocations, aes(x = age, y = total_relocations), color = "grey80", size = 1) +
  geom_segment(data = segment_data, aes(x = age, y = total_relocations, 
                                        xend = xend, yend = yend, color = age_group), size = 1.5) +
  geom_point(data = age_relocations, aes(x = age, y = total_relocations, color = age_group), size = 2) +
  scale_x_continuous(breaks = seq(0, 99, 5)) +
  scale_y_continuous(breaks = seq(0, 20000, 4000)) +
  labs(
    title = "Total Relocations per Age (All Municipalities)",
    x = "Age",
    y = "Total Relocations",
    color = "Age Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank()
  ) + 
  scale_color_manual(values = c(
    "0–17" = "#E41A1C",   # red
    "18–29" = "#377EB8",  # blue
    "30–44" = "#4DAF4A",  # green
    "45–65" = "#984EA3",  # purple
    "66–99" = "#FF7F00"   # orange
  ))
ggsave("out/total_relocations_by_age.png", width = 10, height = 6, dpi = 300)
```


```{r}

population_df <- read_csv('new_data/age_cleaned.csv')

# Prepare a dataframe for storing total relocations per age
age_relocations <- data.frame()

# Loop through each age (0 to 99)
for (age in 0:99) {
  total_relocations <- df %>%
    filter(!is.na(.[[as.character(age)]])) %>%
    summarise(total_relocations = sum(.[[as.character(age)]], na.rm = TRUE)) %>%
    mutate(age = age)
  
  age_relocations <- bind_rows(age_relocations, total_relocations)
}

# Merge with population data and compute relocation rate per 1000 people
age_relocations <- age_relocations %>%
  left_join(population_df, by = "age") %>%
  mutate(
    relocation_rate = (total_relocations / population) * 1000,
    age_group = case_when(
      age <= 17 ~ "0–17",
      age <= 29 ~ "18–29",
      age <= 44 ~ "30–44",
      age <= 65 ~ "45–65",
      TRUE ~ "66–99"
    )
  )

# Create segment data for colored line segments
segment_data <- age_relocations %>%
  arrange(age) %>%
  mutate(
    xend = lead(age),
    yend = lead(relocation_rate),
    group_end = lead(age_group)
  ) %>%
  filter(!is.na(xend))

# Plot: relocation rate per 1000 people
ggplot() +
  geom_line(data = age_relocations, aes(x = age, y = relocation_rate), color = "grey80", size = 1) +
  geom_segment(data = segment_data, aes(x = age, y = relocation_rate,
                                        xend = xend, yend = yend, color = age_group), size = 1.5) +
  geom_point(data = age_relocations, aes(x = age, y = relocation_rate, color = age_group), size = 2) +
  scale_x_continuous(breaks = seq(0, 99, 5)) +
  labs(
    title = "Relocation Rate per Age (per 1000 people)",
    x = "Age",
    y = "Relocations per 1000 People",
    color = "Age Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank()
  ) +
  scale_color_manual(values = c(
    "0–17" = "#E41A1C",   # red
    "18–29" = "#377EB8",  # blue
    "30–44" = "#4DAF4A",  # green
    "45–65" = "#984EA3",  # purple
    "66–99" = "#FF7F00"   # orange
  ))

ggsave("out/relocation_rate_by_age.png", width = 10, height = 6, dpi = 300)


### note: this plot may look very similar to the previous one but this one takes into the skewness of age distribution, i.e. there being fewer 99 year olds than 18 year olds.
```

# creating heatmaps for age groups
```{r}
# # ved ikke hvorfor jeg skal skrive de næste linjer kode, men af en eller grund virkede det ikke når jeg ikke gjorde det
# munic <- gadm(country = "DNK", level = 2, path = ".")
# 
# # Convert to sf
# munic_sf <- st_as_sf(munic)
# 
# # Rename Århus to Aarhus in the NAME_2 column
# munic_sf$NAME_2 <- gsub("^Århus$", "Aarhus", munic_sf$NAME_2)
# munic_sf$NAME_2 <- gsub("^Høje Taastrup$", "Høje-Taastrup", munic_sf$NAME_2)
# munic_sf$NAME_2 <- gsub("^Vesthimmerland$", "Vesthimmerlands", munic_sf$NAME_2)
# 
# # ─── Define age groups ─────────────────────────────────────────────────────────
# age_groups <- list(
#   "0-17"   = 0:17,
#   "18-29"  = 18:29,
#   "30-44"  = 30:44,
#   "45-65"  = 45:65,
#   "66-99"  = 66:99
# )
# 
# # Prepare a dataframe for storing results
# relocations_by_muni <- data.frame()
# 
# # Calculate total relocations by age group and destination
# for (age_label in names(age_groups)) {
#   age_range <- age_groups[[age_label]]
#   age_cols <- as.character(age_range)
#   
#   # Sum all ages in the group
#   relocations <- df %>%
#     mutate(total = rowSums(select(., all_of(age_cols)), na.rm = TRUE)) %>%
#     group_by(dest) %>%
#     summarise(total_relocations = sum(total, na.rm = TRUE)) %>%
#     mutate(age_group = age_label)
#   
#   # Append to the main dataframe
#   relocations_by_muni <- bind_rows(relocations_by_muni, relocations)
# }
# 
# # Join with the population data
# relocations_by_muni <- relocations_by_muni %>%
#   left_join(population_df, by = c("dest" = "municipality")) %>%
#   mutate(relocations_per_capita = total_relocations / as.numeric(population_jan_2025))
# 
# 
# relocations_by_muni <- relocations_by_muni %>% filter(!is.na(relocations_per_capita))
# 
# # Join with the spatial data
# munic_sf <- munic_sf %>%
#   left_join(relocations_by_muni, by = c("NAME_2" = "dest")) %>%
#   filter(!is.na(age_group))
# 
# 
# # Plotting the heatmap
# ggplot(munic_sf) +
#   geom_sf(aes(fill = relocations_per_capita)) +
#   facet_wrap(~age_group) +
#   scale_fill_viridis_c(option = "plasma", trans = "log") +
#   labs(
#     title = "Per Capita Relocations to Each Municipality by Age Group",
#     fill = "Relocations per Capita"
#   ) +
#   theme_minimal(base_size = 14) +
#   theme(
#     legend.position = "right",
#     strip.text = element_text(size = 12, face = "bold")
#   )

```

### creating cartograms
```{r}
munic <- gadm(country = "DNK", level = 2, path = ".")

# Convert to sf
munic_sf <- st_as_sf(munic)

# Rename Århus to Aarhus in the NAME_2 column
munic_sf$NAME_2 <- gsub("^Århus$", "Aarhus", munic_sf$NAME_2)
munic_sf$NAME_2 <- gsub("^Høje Taastrup$", "Høje-Taastrup", munic_sf$NAME_2)
munic_sf$NAME_2 <- gsub("^Vesthimmerland$", "Vesthimmerlands", munic_sf$NAME_2)

# Transform to a Projected Coordinate System (ETRS89 / UTM zone 32N)
munic_sf <- st_transform(munic_sf, crs = 25832)

# ─── Define age groups ─────────────────────────────────────────────────────────
age_groups <- list(
  "0-17"   = 0:17,
  "18-29"  = 18:29,
  "30-44"  = 30:44,
  "45-65"  = 45:65,
  "66-99"  = 66:99
)

# Loop Over Age Groups
for (group_name in names(age_groups)) {
  
  # Get the Age Range
  age_range <- age_groups[[group_name]]
  
  # Calculate Total Relocations per Municipality for the Current Age Group
  relocations_per_muni <- df %>%
    group_by(dest) %>%
    summarise(total_relocations = sum(across(as.character(age_range)), na.rm = TRUE)) %>%
    ungroup()
  
  # Join with Spatial Data
  muni_group_sf <- munic_sf %>%
    left_join(relocations_per_muni, by = c("NAME_2" = "dest")) %>%
    filter(!is.na(total_relocations))
  
  # Create the Cartogram
  muni_cartogram <- cartogram_cont(muni_group_sf, weight = "total_relocations", itermax = 10)
  
  # Plot the Cartogram
  p <- ggplot(muni_cartogram) +
    geom_sf(aes(fill = total_relocations), color = "black", size = 0.2) +
    scale_fill_viridis_c(option = "plasma", trans = "log", limits = c(1, 30000)) +
    labs(
      title = paste("Cartogram of Relocations (Age", group_name, ") per Municipality in Denmark"),
      fill = "Total Relocations"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position = "right",
      plot.title = element_text(size = 16, face = "bold")
    )
  
  # Save the Plot
  ggsave(filename = paste0("out/relocations_wee__cartogram_", group_name, ".png"), plot = p, width = 10, height = 8, dpi = 300)
}

```

```{r}
pacman::p_load('magick')

# Define your age labels (in the same order as your cartogram PNGs)
age_labels <- c("0-17", "18-29", "30-44", "45-65", "66-99")
output_dir <- "out"
gif_dir <- "gifs"
dir.create(gif_dir, showWarnings = FALSE)

# Generate file paths for the images
image_files <- paste0(output_dir, "/relocations_wee__cartogram_", age_labels, ".png")

# Read and animate the images
if (all(file.exists(image_files))) {
  images <- image_read(image_files)
  animation <- image_animate(image_join(images), fps = 0.5)  # 1 frame every 4 seconds

  # Save the GIF
  gif_filename <- file.path(gif_dir, "relocation_cartogram.gif")
  image_write(animation, gif_filename)
  message("GIF saved to: ", gif_filename)
} else {
  warning("One or more image files are missing.")
}

```

### creating popularity cartograms
```{r}
library(readxl)

# Replace with your actual file path
population_file_path <- "new_data/FOLK1AM.xlsx"

# Extract the relevant data
munic_pop_df <- read_excel(population_file_path, skip = 2) %>%
  select(municipality = `...1`,
         population_jan_2025 = `2024M12`) %>%
  mutate(municipality = trimws(municipality)) %>%
  filter(!is.na(municipality), municipality != "Hele landet") %>%
  filter(!municipality %in% c("Region Hovedstaden", "Region Sjælland", 
                              "Region Syddanmark", "Region Midtjylland", 
                              "Region Nordjylland"))

# View the cleaned population data
head(munic_pop_df)
```

```{r}

# ─── Load Municipalities ───────────────────────────────────────────────
munic <- gadm(country = "DNK", level = 2, path = ".")

munic_sf <- st_as_sf(munic)
munic_sf$NAME_2 <- gsub("^Århus$", "Aarhus", munic_sf$NAME_2)
munic_sf$NAME_2 <- gsub("^Høje Taastrup$", "Høje-Taastrup", munic_sf$NAME_2)
munic_sf$NAME_2 <- gsub("^Vesthimmerland$", "Vesthimmerlands", munic_sf$NAME_2)
munic_sf <- st_transform(munic_sf, crs = 25832)

# ─── Load Population Data ──────────────────────────────────────────────
population_file_path <- "/Users/villiamjensen/Downloads/FOLK1AM.xlsx"

munic_pop_df <- read_excel(population_file_path, skip = 2) %>%
  select(municipality = `...1`,
         population_jan_2025 = `2024M12`) %>%
  mutate(municipality = trimws(municipality)) %>%
  filter(!is.na(municipality), municipality != "Hele landet") %>%
  filter(!municipality %in% c("Region Hovedstaden", "Region Sjælland",
                              "Region Syddanmark", "Region Midtjylland",
                              "Region Nordjylland"))

# ─── Clean df$dest to match municipality names ─────────────────────────
df$dest <- gsub("^Århus$", "Aarhus", df$dest)
df$dest <- gsub("^Høje Taastrup$", "Høje-Taastrup", df$dest)
df$dest <- gsub("^Vesthimmerland$", "Vesthimmerlands", df$dest)
df$dest <- trimws(df$dest)

# ─── Define Age Groups ─────────────────────────────────────────────────
age_groups <- list(
  "0-17"   = 0:17,
  "18-29"  = 18:29,
  "30-44"  = 30:44,
  "45-65"  = 45:65,
  "66-99"  = 66:99
)

# ─── Precompute Popularity Indices ─────────────────────────────────────
popularity_all <- map_dfr(names(age_groups), function(group_name) {
  age_range <- age_groups[[group_name]]
  
  relocations_per_muni <- df %>%
    group_by(dest) %>%
    summarise(total_relocations = sum(across(as.character(age_range)), na.rm = TRUE)) %>%
    ungroup()
  
  relocations_per_muni %>%
    left_join(munic_pop_df, by = c("dest" = "municipality")) %>%
    mutate(
      popularity_index = (total_relocations / population_jan_2025) * 1000,
      age_group = group_name
    ) %>%
    filter(!is.na(popularity_index))
})

# ─── Plot Cartograms ───────────────────────────────────────────────────
for (group_name in names(age_groups)) {
  
  group_data <- popularity_all %>%
    filter(age_group == group_name)
  
  muni_group_sf <- munic_sf %>%
    left_join(group_data, by = c("NAME_2" = "dest")) %>%
    filter(!is.na(popularity_index))
  
  muni_cartogram <- cartogram_cont(muni_group_sf, weight = "popularity_index", itermax = 10)
  
  p <- ggplot(muni_cartogram) +
    geom_sf(aes(fill = popularity_index), color = "black", size = 0.2) +
    scale_fill_viridis_c(option = "plasma", trans = "log", limits = c(1, 65)) + ##the values were found by summary(popularity_all$popularity_index)
    labs(
      title = paste("Popularity Index (Age",group_name,"): Relocations per 1000 Residents"),
      fill = "Relocations per 1000 Residents"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position = "right",
      plot.title = element_text(size = 16, face = "bold")
    )
  
  ggsave(filename = paste0("out/popularity_cartogram_", group_name, ".png"),
         plot = p, width = 10, height = 8, dpi = 300)
}

# ─── Create Animated GIF ───────────────────────────────────────────────
pacman::p_load('magick')

age_labels <- names(age_groups)
output_dir <- "out"
gif_dir <- "gifs"
dir.create(gif_dir, showWarnings = FALSE)

image_files <- paste0(output_dir, "/popularity_cartogram_", age_labels, ".png")

if (all(file.exists(image_files))) {
  images <- image_read(image_files)
  animation <- image_animate(image_join(images), fps = 0.5)
  
  gif_filename <- file.path(gif_dir, "popularity_cartogram.gif")
  image_write(animation, gif_filename)
  message("GIF saved to: ", gif_filename)
} else {
  warning(" One or more image files are missing.")
}


```


